---
title: "EFS Survival Analysis \n COG meeting Spring 2017 "
author: "Jenny Smith"
date: "April 25, 2017"
output: pdf_document
---

Purpose: To create EFS curves to identify if new CKIT patients or MYC-ITD which are associated with patient outcome. 

```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = 'H:/COG_Meetings_Analysis/2017.04.21_TARGET_AML_cKITpatients/')
```


```{r message = FALSE, warning=FALSE}
library(survival)
library(RColorBrewer)
library(colorspace)
library(stringr)
library(Biobase)
library(dplyr)
library(magrittr)
library(tables)
getwd()
```


```{r}
source("H:/scripts/Survival_Analysis/Survplot_Functions.r")
```


#Read in the clinical data with the new classification strategy
```{r comment="", tidy=TRUE, size="tiny"}
# newClassification <- read.csv("New_stratification_outcomes_updated)_(003).csv")

newClassification <- read.csv("H:/reference_mapping-files/New_stratification_outcomes_updated_2017.03.28.csv", stringsAsFactors = FALSE)

head(newClassification[,c(1,9:12)])
```

```{r}
dim(newClassification)
```


#Combine the new C-Kit Mutations to see if we add any new Patiets.

```{r}
kit <- read.csv("cKit_Allmutations.csv", stringsAsFactors = FALSE)
head(kit[,1:6]) #note many repeated patient USIs 
```

```{r}
#Identify which KITs are already in the new classification scheme
KIT_newClass <- newClassification[which(grepl("KIT",  newClassification$Reason)), ]$TARGET.USI
length(which(! kit$patientID %in%  KIT_newClass )) #22 not included already
```

```{r}
#must remove duplicates or it will be duplicated in the merge.
kit <- kit[which(!kit$patientID %in% KIT_newClass), ]
kit <- kit[which(! duplicated(kit$patientID)), ]  

dim(kit)
```

```{r}
newClassUpdate <- merge(kit[,1:2], newClassification, by.x = "patientID", by.y="TARGET.USI",  all.y=TRUE)

newClassUpdate[90:100,c(1:2,11)]
```

```{r}
#if "gene" column is.na() then leave reason as is. else add +KIT to the string. 
newClassUpdate$Reason <- ifelse(is.na(newClassUpdate$gene), newClassUpdate$Reason, sprintf("%s + KIT", newClassUpdate$Reason))

#remove the "gene" column 
newClassUpdate <- newClassUpdate[,-2] 

newClassUpdate[90:100,c(1,10)]
```

#Add MYC-ITD to New Stratification

```{r}
MYC_ITD <- read.csv("H:/COG_Meetings_Analysis/2017.04.21_TARGET_AML_cKITpatients/MYC_ITD_0531_prevalence _21216.csv", stringsAsFactors = FALSE)
```

```{r}
newClassUpdate <- merge(MYC_ITD[, c("Reg.", "MYC.TD.by.DNA")], newClassUpdate, by.x = "Reg.", by.y = "Reg.", all.y = TRUE)

dim(newClassUpdate)
```

```{r}
newClassUpdate$MYC_ITD <- ifelse(is.na(newClassUpdate$MYC.TD.by.DNA), "Unknown", newClassUpdate$MYC.TD.by.DNA ) %>% { gsub("yes.+", "Yes", . ) } %>% { gsub("^no", "No", . ) }
```


```{r}
newClassUpdate <- newClassUpdate[, -2] #remove the extra column 
```


```{r}
newClassUpdate[90:100,c(2,10,82)]
```

#Save the new classifications update

```{r}
write.csv(newClassUpdate, file="New_stratification_outcomes_updatedbyJS_2017.04.25.csv", row.names = FALSE)
```



#Define Functions for subsetting and KM curves

##Customize the output of SurvivalPlot() function. 
SurvivalPlot is a function that is my generic version in Survplot_Functions.r
```{r}
CustomPlot <- function(fit, diff, color){
    
    #label with number of patients
    if (length(fit$n) == 1){
      Num <- paste("n =", fit$n, sep = " ")
    }else{
      obs <- fit$n[1] + fit$n[2]
      Num <- paste("n =", obs, sep = " ")
    }
    
    #X axis position for Num
    pos <- (max(fit$time)*0.75)
    
    #label with pvalue
    pVal <- 1-pchisq(diff$chisq, (length(diff$n)-1))
    pVal <- round(pVal, digits = 4)
    
    if (pVal < 0.001){
      p <- "p < 0.001"
    }else{
      p <- paste("p =", pVal, sep = " ")
    }
    
    #customized plot 
    if (length(color > 1)){
      efs_plot <- SurvivalPlot(fit, "Mutation Status", "Years",color)
    }else{
      efs_plot <- SurvivalPlot(fit, "NA", "Years",color)
    }

    efs_plot <- efs_plot + annotate(geom="text", x=pos, y=1.0, label= Num, size=6)
    
    #return the ggplot object
    return(efs_plot)
}
```


#Kaplan-Meier Curves of EFS (Event Free Survival)

#Survival Analysis for MYC-ITD with known status as present or absent. 

```{r}
knownMYCStatus <- newClassUpdate[which(grepl("Yes|No", newClassUpdate$MYC_ITD)), ]
```


```{r}
# SA_list <- list()
# for (type in c("EFS", "OS")){
#   SA <- SurvObjects(knownMYCStatus, "MYC_ITD", 1, type)
#   plot <- efsCustomPlot(SA[[1]], c("royalblue", "darkorchid"))
#   title <- paste(c(type, "in Patients with MYC-ITD"), sep=" ")
#   plot$labels$title <- title
}
```


```{r}
MYC_SA <- SurvObjects(knownMYCStatus, "MYC_ITD", 1, "EFS")

MYC_SA
```

```{r}
MYC_EFS <- efsCustomPlot(MYC_SA[[1]], c("royalblue", "darkorchid"))

MYC_EFS$labels$title <- "Event Free Survival in Patients with MYC-ITD"

MYC_EFS <- MYC_EFS + annotate("text", x=0.5, y=0.08, label= c("p = 0.155"))

MYC_EFS
```

#MYC-ITD Present Against all others
```{r}
MYC_Vs_All <- newClassUpdate
MYC_Vs_All$MYC_ITD <- ifelse(MYC_Vs_All$MYC_ITD == "Yes", MYC_Vs_All$MYC_ITD, "All Others" )
```

```{r}
MYC_vsAll_SA <-  SurvObjects(MYC_Vs_All, "MYC_ITD", 1, "EFS")

MYC_vsAll_SA
```

```{r}
efsCustomPlot(MYC_vsAll_SA[[1]], c("royalblue", "darkorchid")) + annotate("text", x=0.5, y=0.08, label= c("p = 0.099 ")) + labs(title="Event Free Survival in Patients with MYC-ITD")
```

#cKit Mutations efs

```{r}
group1 <- newClassUpdate[which(grepl("RUNX1-RUNX1T1 \\+ KIT", newClassUpdate$Reason) | grepl("CBFB-MYH11 \\+ KIT", newClassUpdate$Reason)), ]

CBF_KIT <- rep("CBF and cKIT", nrow(group1))

group1 <- cbind(CBF_KIT, group1)
```

```{r}
group2 <- newClassUpdate[which(grepl("RUNX1-RUNX1T1", newClassUpdate$Reason) | grepl("CBFB-MYH11", newClassUpdate$Reason)), ]

group2 <- group2[ ! group2$TARGET.USI %in% group1$TARGET.USI, ]

CBF_KIT <- rep("CBF Alone", nrow(group2))

group2 <- cbind(CBF_KIT, group2)
```

```{r}
dim(group1)
dim(group2)
```


```{r}
KIT_df <- rbind(group1,group2)

head(KIT_df[,c(1,9:12)])
```

### Survival analysis
```{r message=FALSE, warning=FALSE}
KIT_SA <- SurvObjects(KIT_df, "CBF_KIT", 1, "EFS")

KIT_SA
```


### KM plots

```{r message=FALSE, warning=FALSE}
# KIT <- SurvivalPlot(KIT_SA[[1]], "Mutation Status", "Year", c("dodgerblue4","darkturquoise"))
KIT <- efsCustomPlot(KIT_SA[[1]],  c("dodgerblue4","darkturquoise"))
```

```{r}
KIT$labels$title <- "Event Free Survival in \n Patients with CBF and cKIT Mutations"

KIT
```





